---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ost.utils

<!-- badges: start -->
[![R-CMD-check](https://github.com/pedrobsantos21/ost.utils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pedrobsantos21/ost.utils/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `ost.utils` package provides utility methods and tools to ease the workflow in projects developed by Coordenadoria do Observatório de Segurança no Trânsito (COST) of Detran-SP.

## Installation

You can install the `ost.utils` from GitHub with:

```r
# install.packages("pak")
pak::pak("pedrobsantos21/ost.utils")
```

## Example

This is a basic example which shows you how to download and clean Infosiga.SP data:

```{r}
library(ost.utils)


df_sinistros <- load_infosiga(
  file_type = "sinistros", 
  zip_path = "data-raw/dados_infosiga.zip"
)

df_sinistros_clean <- clean_infosiga(df_sinistros, "sinistros")

head(df_sinistros_clean)
```

## License

This package is licensed under the GPL (>= 3) license.

## Citation

```{r citation}
citation("ost.utils")
```
