---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ostutils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `ost.utils` package provides tools and methods to streamline the workflow in R projects developed by the Coordenadoria do Observatório de Segurança no Trânsito (COST) of Detran-SP.

## Installation

The development version of `ost.utils` can be installed from [GitHub](https://github.com/pedrobsantos21/ost.utils) with:

```r
# install.packages("pak")
pak::pak("pedrobsantos21/ost.utils")
```

## Package organization

This package is organized into two main groups of functions:

1. `infosiga`: methods to download, load and clean open data from [Infosiga.SP](https://infosiga.detran.sp.gov.br)

- `download_infosiga()`

- `load_infosiga()`

- `clean_infosiga()`

2. `plot`: helper functions to plot data with `ggplot2`:

- `palette_detran()`

- `theme_detran()`

## Usage example

In this example, we will load Infosiga road crash data and plot it using `ggplot2`. First, we load the required packages:

```{r setup}
library(ost.utils)
library(ggplot2)
library(dplyr)
library(lubridate)
```

Then, we use `download_infosiga()` to save the data to a temporary folder, load the road crash data with `load_infosiga()`, and clean it with `clean_infosiga()`. In a typical project, you might download the data to a dedicated `data/` folder.

```{r}
temp <- tempdir()
download_infosiga(temp)

df <- load_infosiga(file_type = "sinistros", path = temp)

df_clean <- clean_infosiga(df, file_type = "sinistros")

head(df_clean)
```

Now we can plot the count of road crashes per year using the custom Detran style:

```{r}
df_clean |> 
  filter(
    tipo_registro %in% c("Sinistro fatal", "Sinistro não fatal"),
    year(data_sinistro) > 2018
  ) |> 
  count(year = year(data_sinistro)) |> 
  ggplot(aes(x = year, y=n)) +
  geom_col(fill = palette_detran()$blue) +
  theme_detran()

```
